SubsidyExplorer - Step 3: Model Parameterization

### 2020-08-07 scripts/04_model_parameterization.Rmd
model param updated

This script handles the preliminary data wrangling necessary to create the biological parameters for the bio-economic model underlying the SubsidyExplorer toolkit. It outputs two CSV files used by the SubsidyExplorer toolkit (located in `r here::here("WTOToolkit", "data")`): 
- model_parameters_global.csv
- model_parameters_regional.csv

@@

# Path to vessel list
vessel_list_path <- here::here("results", "03-management", "vessel_list_2018_management.csv")
```

In addition to the data sources described in the inital data wrangling script, we also draw on the folloing data sources:

```{r setup, include=FALSE}
# Code options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = F)

@@

# Define regions 
region_names <- sort(c("atlantic_e", "atlantic_w", "indian", "pacific_e", "pacific_w", "polar"))

@@

regional_model_parameters <- tibble(parameter = character(0), source = character(0)) %>%
  bind_cols(regions_spread)
```

@@

We assume $\phi = 1.188$ based on literature for all (Thorson et al. 2012; World Bank 2017).

@@

```{r}
# One estimate of global carrying capacity
global_k <- sum(costello_2016_dat$k)/0.78
#global_k <- global_msy*9.1

@@

# Add g to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "g", value = global_g, source = "Costello et al. 2016"))

@@

# Add g to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "r", value = global_r, source = "Costello et al. 2016"))
```

@@

$$ h_0+ \xi =\frac{\phi + 1}{\phi} r b_{0}\Bigg(1 - \bigg(\frac{b_{0}}{K}\bigg)^{\phi}\Bigg)$$

@@

# Add base year harvest to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "h0", value = global_h0, source = "FAO 2019"))
```

```{r}
# Estimation of starting biomass
biomass_vec <- seq(0, 0.5*global_k, length.out = 10000)
growth_approx <- (((phi+1)/phi)*global_g*biomass_vec*(1-((biomass_vec/global_k)^phi)))
biomass_pos <- which.min(abs(growth_approx - (global_h0+8000000)))
global_b0 <- biomass_vec[biomass_pos]

@@

We also apply our bio-economic model of fisheries subsidies reform on a regional basis. For this analysis, we have chosen to use six regions based loosly on the major ocean basins and using the FAO statistical regions as subdivisions as needed to further separate regions: Atlantic Ocean (East) and Mediterranean [`r paste(regions_tidy$fao_region[regions_tidy$region == "atlantic_e"], collapse = ", ")`], Atlantic Ocean (West) [`r paste(regions_tidy$fao_region[regions_tidy$region == "atlantic_w"], collapse = ", ")`], Indian Ocean [`r paste(regions_tidy$fao_region[regions_tidy$region == "indian"], collapse = ", ")`], Pacific Ocean (East) [`r paste(regions_tidy$fao_region[regions_tidy$region == "pacific_e"], collapse = ", ")`], Pacific Ocean (West) [`r paste(regions_tidy$fao_region[regions_tidy$region == "pacific_w"], collapse = ", ")`], and Polar Oceans [`r paste(regions_tidy$fao_region[regions_tidy$region == "polar"], collapse = ", ")`].

@@

```{r}
# Extract carrying capacity
#regional_k <- regional_msy*9.1
regional_k <- round(costello_2016_dat_regional$k/0.78)
names(regional_k) <- region_names

@@

# Add r to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "g", "source" = "Costello et al. 2016", regional_g))

@@

# Add r to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "r", "source" = "Costello et al. 2016", regional_r))
```

@@

```{r}
#Estimate starting biomass
#regional_b0 <- numeric(length(region_names))
# regional_biomass_growth <- c(1.33e6, 1.33e6, 1.33e6, 1.33e6, 1.33e6, 1.33e6)
# 
# for(i in 1:length(regional_h0)){
#   biomass_vec <- seq(0, regional_k[i], length.out = 10000)
#   growth_approx <- (((regional_phi[i]+1)/regional_phi[i])*regional_r[i]*biomass_vec*(1-((biomass_vec/regional_k[i])^regional_phi[i])))
#   biomass_pos <- which.min(abs(growth_approx - (regional_h0 + regional_biomass_growth[i])))
#   regional_b0[i] <- biomass_vec[biomass_pos] 
# }

regional_b0 <- regional_k*0.2
names(regional_b0) <- region_names

@@

<!-- We can then calculate the price constant using 2012 data from Costello et al. (2016).  -->

```{r}
# Global price constant
global_price_wrangling <- costello_2016_dat %>%
  group_by(year) %>%
  summarize(price = median(price, na.rm = T),
            harvest = sum(catch, na.rm = T))

global_delta <- global_price_wrangling$harvest/(global_price_wrangling$price^global_epsilon)

@@

<!-- We then calculate the price constant for each region using 2012 data from Costello et al. (2016).  -->

```{r}
# Regional price constant
regional_delta <- costello_2016_dat_regional$catch/(costello_2016_dat_regional$price^regional_epsilon)

@@

In order to calculate $\alpha$, we first need to determine fish price in the base year. We compare two alternatives - 1) using the median price from Costello et al. (2016) adjusted to 2018 US$, and 2) solving for price based on the observed base year harvest.

```{r}
# Calculate global base prices
global_base_price1 <- ((1/global_delta)^(1/global_epsilon))*(global_h0^(1/global_epsilon))

# CPI_2012 <- 229.594  #BLS CPI All Items (Annual Ave), (BLS Series ID: CUUR0000SA0)
# CPI_2018 <- 251.107  #BLS CPI All Items (Annual Ave), (BLS Series ID: CUUR0000SA0)
# 
# global_base_price2 <- global_price_wrangling$price * (CPI_2018/CPI_2012)
```

The second price looks a bit high (comparable global price estimates are in the range of $1200-1400 per ton), so we use the first.

```{r}
# Add to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "p0", value = global_base_price1, source = "Costello et al. 2016"))
```

@@

# Add to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "eta", value = global_eta, source = "Calculated"))

@@

We then calculate price in the base year in the same way regionally

```{r}
# Calculate base price regionally
regional_base_price1 <- ((1/regional_delta)^(1/regional_epsilon))*(regional_fleet$catch^(1/regional_epsilon))

# Add catchability to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "p0", "source" = "Calculated", regional_base_price1))

@@

@@

```{r}
regional_omega <- rep(0.0011, length = length(region_names))
names(regional_omega) <- region_names

@@

write_csv(regional_model_parameters, paste0(results_dir, "model_parameters_regional.csv"))
write_csv(regional_model_parameters, paste0(app_dir, "model_parameters_regional.csv"))
```

>>>>

This script handles the preliminary data wrangling necessary to create the biological parameters for the bio-economic model underlying the SubsidyExplorer toolkit. It allows for specification of different regional configurations, though we have selected the one for which we believe the data is most supportive.

@@

# Path to vessel list
vessel_list_path <- here::here("results", "03-management", "vessel_list_2018_management.csv")

# Code options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = F)

@@

# Select number of regions (choices are 3 or 6)
n_regions <- 3

In addition to the data sources described in the inital data wrangling script, we also draw on the folloing sources:

```{r}
if(n_regions == 3){

  # Define regions 
  region_names <- sort(c("atlantic", "indian", "pacific"))

  # Create a tidy df with the cooresponding fao regions
  regions_tidy <- tibble(
    region = region_names,
    fao_region = list(c(21, 27, 31, 37, 34, 41, 47, 48), # atlantic (and med)
                      c(51, 57, 58), # indian
                      c(18, 61, 67, 71, 77, 81, 87, 88)) # pacific 
  ) %>%
    unnest(cols = c(fao_region))

  # Also create a blank tibble with regions as column names
  regions_spread <- setNames(data.frame(matrix(ncol = length(region_names), nrow = 0)), as.character(region_names))

  regional_model_parameters <- tibble(parameter = character(0), source = character(0)) %>%
    bind_cols(regions_spread) %>%
    mutate(atlantic = as.character(atlantic),
           indian = as.character(indian),
           pacific = as.character(pacific))
}
```

```{r setup, include=FALSE}
if(n_regions == 6){

# Define regions 
region_names <- sort(c("atlantic_e", "atlantic_w", "indian", "pacific_e", "pacific_w", "polar")) # ran into difficulties with the polar region. 6 regions needs more work

@@

regional_model_parameters <- tibble(parameter = character(0), source = character(0)) %>%
  bind_cols(regions_spread)

}
```

@@

We assume $\phi = 1.188$ based on literature for all (Thorson et al. 2012; World Bank 2017; Costello et al. 2016).

@@

```{r}
# One estimate of global carrying capacity
global_k <- global_msy*9.1

@@

# Add g to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "g", value = global_g, source = "Calculated"))

@@

# Add g to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "r", value = global_r, source = "Calculated"))
```

@@

$$ h_0+ \xi =\frac{\phi + 1}{\phi} g b_{0}\Bigg(1 - \bigg(\frac{b_{0}}{K}\bigg)^{\phi}\Bigg)$$

@@

# Add base year harvest to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "h0", value = global_h0, source = "FAO 2020"))
```

```{r}
# Estimation of starting biomass
biomass_vec <- seq(0, 0.75*global_k, length.out = 1000000)
growth_approx <- (((phi+1)/phi)*global_g*biomass_vec*(1-((biomass_vec/global_k)^phi)))
biomass_pos <- which.min(abs(growth_approx - (global_h0+8000000)))
global_b0 <- biomass_vec[biomass_pos]

@@

We also apply our bio-economic model of fisheries subsidies reform on a regional basis.

@@

```{r}
# Extract carrying capacity
regional_k <- regional_msy*9.1

@@

# Add r to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "g", "source" = "Calculated", regional_g))

@@

# Add r to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "r", "source" = "Calculated", regional_r))
```

@@

We assume biomass growth to be equally divided among regions.

```{r}
#Estimate starting biomass
regional_b0 <- numeric(length(region_names))
regional_biomass_growth <- c(2.66e6, 2.66e6, 2.66e6)

for(i in 1:length(regional_h0)){
  biomass_vec <- seq(0, 0.75*regional_k[i], length.out = 10000)
  growth_approx <- (((regional_phi[i]+1)/regional_phi[i])*regional_g[i]*biomass_vec*(1-((biomass_vec/regional_k[i])^regional_phi[i])))
  biomass_pos <- which.min(abs(growth_approx - (regional_h0 + regional_biomass_growth[i])))
  regional_b0[i] <- biomass_vec[biomass_pos]
}

names(regional_b0) <- region_names

@@

@@

@@

In order to calculate $\alpha$, we first need to determine fish price in the base year. We compare two alternatives - 1) using the median price from Costello et al. (2016) adjusted to 2018 US\$, and 2) using the global price from the latest Sunken Billions report adjusted to 2018 US\$. The first price looks a bit high (comparable global price estimates are in the range of $1200-1400 per ton), so we use the second

```{r}
CPI_2012 <- 229.594  #BLS CPI All Items (Annual Ave), (BLS Series ID: CUUR0000SA0)
CPI_2018 <- 251.107  #BLS CPI All Items (Annual Ave), (BLS Series ID: CUUR0000SA0)

# Global price constant
global_price_wrangling <- costello_2016_dat %>%
  group_by(year) %>%
  summarize(price = median(price, na.rm = T))

global_base_price1 <- global_price_wrangling$price*(CPI_2018/CPI_2012)
global_base_price2 <- 1260*(CPI_2018/CPI_2012)

# Add to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "p0", value = global_base_price2, source = "World Bank 2017"))
```

```{r}
global_delta <- global_h0/(global_base_price2^global_epsilon)

@@

# Add to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "eta", value = global_eta, source = "Calculated"))
```

@@

For simplicity, we assume the starting price to be equal across all regions.

```{r}
# Calculate base price regionally
regional_base_price <- rep(global_base_price2, length = length(region_names))
names(regional_base_price) <- region_names

# Add catchability to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "p0", "source" = "World Bank 2017", regional_base_price))

@@

```{r}
# Regional price constant
regional_delta <- regional_h0/(regional_base_price^regional_epsilon)

@@

```{r}
regional_omega <- rep(0.001, length = length(region_names))
names(regional_omega) <- region_names

@@

regional_name <- paste0("model_parameters_regional_", n_regions, "_regions.csv")

write_csv(regional_model_parameters, paste0(results_dir, regional_name))
write_csv(regional_model_parameters, paste0(app_dir, regional_name))
```
